---
title: "Assignment 2 version 2"
author: "Rowan Muhl"
date: "2023-10-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### PART ONE ####


```{r}
#| message = FALSE
library(tidyverse)
library(gtrendsR)
library(censusapi)
library(dplyr)
```


```{r}
res_hw1 <- gtrends(c("crime", "loans"), 
               geo = "US-IL", 
               time = "2020-01-01 2020-12-31", 
               low_search_volume = TRUE)
plot(res_hw1)
```
#### Question 1 ####

Separate out the data on each keyword
```{r}
crime_data <- res_hw1$interest_over_time[which(res_hw1$interest_over_time$keyword == "crime"), ]
loans_data <- res_hw1$interest_over_time[which(res_hw1$interest_over_time$keyword == "loans"), ]
```

Calculate the mean for both keywords
```{r}
mean_crime <- mean(crime_data$hits)
mean_crime
mean_loans <- mean(loans_data$hits)
mean_loans
```

Calculate the median for both keywords
```{r}
median_crime <- median(crime_data$hits)
median_crime
median_loans <- median(loans_data$hits)
median_loans
```

Calculate the variance for both keywords
```{r}
variance_crime <- var(crime_data$hits)
variance_crime
variance_loans <- var(loans_data$hits)
variance_loans
```

Make a function that will calculate the mode
```{r}
Mode <- function(x) {
  uniq_x <- unique(x)
  uniq_x[which.max(tabulate(match(x, uniq_x)))]
}
```

Calculate the mode for both keywords
```{r}
mode_crime <- Mode(crime_data$hits)
mode_crime
mode_loans <- Mode(loans_data$hits)
mode_loans
```

#### Question 2 ####

focus on the interest_by_city part
```{r}
loans_data2 <- res_hw1$interest_by_city[which(res_hw1$interest_by_city$keyword == "loans"), ]

```

Group the data about the "loans" keyword by city
Then look at the top cities
```{r}
city_search_frequency <- loans_data2 %>%
  group_by(location) %>%
  summarize(total_hits = sum(hits, na.rm = TRUE)) %>%
  arrange(desc(total_hits))

city_search_frequency
```

#### Question 3 ####

Start with a correlation coefficient
```{r}
correlation_coefficient_crimeloans <- cor(crime_data$hits, loans_data$hits, use = "complete.obs")
correlation_coefficient_crimeloans
```

Do a significance test, then check if the correlation is statistically significant
```{r}
cor_test_result <- cor.test(crime_data$hits, loans_data$hits, method = "pearson")
cat("p-value for Correlation Test:", cor_test_result$p.value, "\n")

if (cor_test_result$p.value < 0.05) {
  cat("There is a statistically significant relationship between 'crime' and 'loans'.\n")
} else {
  cat("There is no statistically significant relationship between 'crime' and 'loans'.\n")
}
```


#### Question 4 ####
Now we do everything again but for covid related keywords.
I chose the words "mask" and "symptoms"

```{r}
res_hw2 <- gtrends(c("mask", "symptoms"), 
               geo = "US-IL", 
               time = "2020-01-01 2020-12-31", 
               low_search_volume = TRUE)
plot(res_hw2)
```

Separate out the data on each keyword
```{r}
mask_data <- res_hw2$interest_over_time[which(res_hw2$interest_over_time$keyword == "mask"), ]
symptoms_data <- res_hw2$interest_over_time[which(res_hw2$interest_over_time$keyword == "symptoms"), ]
```

Calculate the mean for both keywords
```{r}
mean_mask <- mean(mask_data$hits)
mean_mask
mean_symptoms <- mean(symptoms_data$hits)
mean_symptoms
```

Calculate the median for both keywords
```{r}
median_mask <- median(mask_data$hits)
median_mask
median_symptoms <- median(symptoms_data$hits)
median_symptoms
```

Calculate the variance for both keywords
```{r}
variance_mask <- var(mask_data$hits)
variance_mask
variance_symptoms <- var(symptoms_data$hits)
variance_symptoms
```

Calculate the mode for both keywords (use the function from earlier)
```{r}
mode_mask <- Mode(mask_data$hits)
mode_mask
mode_symptoms <- Mode(symptoms_data$hits)
mode_symptoms
```

#### Question 5 part 1 ####

focus on the interest_by_city part
```{r}
mask_data2 <- res_hw2$interest_by_city[which(res_hw2$interest_by_city$keyword == "mask"), ]
```

Group the data about the "mask" keyword by city
Then look at the top cities
```{r}
city_search_frequency_mask <- mask_data2 %>%
  group_by(location) %>%
  summarize(total_hits = sum(hits, na.rm = TRUE)) %>%
  arrange(desc(total_hits))

city_search_frequency_mask
```


#### Question 5 part 2 ####

focus on the interest_by_city part
```{r}
symptoms_data2 <- res_hw2$interest_by_city[which(res_hw2$interest_by_city$keyword == "symptoms"), ]
```

Group the data about the "symptoms" keyword by city
Then look at the top cities
```{r}
city_search_frequency_symptoms <- symptoms_data2 %>%
  group_by(location) %>%
  summarize(total_hits = sum(hits, na.rm = TRUE)) %>%
  arrange(desc(total_hits))

city_search_frequency_symptoms
```








#### Question 6 ####

Start with a correlation coefficient
```{r}
correlation_coefficient_masksymptoms <- cor(mask_data$hits, symptoms_data$hits, use = "complete.obs")
correlation_coefficient_masksymptoms
```

Do a significance test, then check if the correlation is statistically significant
```{r}
cor_test_result2 <- cor.test(mask_data$hits, symptoms_data$hits, method = "pearson")
cat("p-value for Correlation Test:", cor_test_result2$p.value, "\n")

if (cor_test_result2$p.value < 0.05) {
  cat("There is a statistically significant relationship between 'mask' and 'symptoms'.\n")
} else {
  cat("There is no statistically significant relationship between 'mask' and 'symptoms'.\n")
}
```


#### PART TWO ####

```{r}
cs_key <- "9298feb975ef6405fa0fc96e9ca3ec7e36bc249d"
acs_il <- getCensus(name = "acs/acs5",
                    vintage = 2020, 
                    vars = c("NAME", 
                             "B01001_001E", 
                             "B06002_001E", 
                             "B19013_001E", 
                             "B19301_001E"), 
                    region = "place:*", 
                    regionin = "state:17",
                    key = cs_key)
head(acs_il)
```

convert missing values to NAs
```{r}
acs_il[acs_il == -666666666] <- NA
```


Now, it might be useful to rename the socio-demographic variables (B01001_001E etc.) in our data set and assign more meaningful names.
```{r}

acs_il <-
  acs_il %>%
  rename(pop = B01001_001E, 
         age = B06002_001E, 
         hh_income = B19013_001E, 
         income = B19301_001E)

```


Add a new variable "location" to the ACS data that only includes city names.

```{r}
no_village <- gsub(' village, Illinois', '', acs_il$NAME)
no_city <- gsub(' city, Illinois', '', no_village)
acs_with_location <- acs_il %>% mutate(location = no_city)
acs_with_location %>% head(5)

```

focus on the interest_by_city part
```{r}
#loans_data2 <- res_hw1$interest_by_city[which(res_hw1$interest_by_city$keyword == "loans"), ]

```

Group the data about the "loans" keyword by city
Then look at the top cities
```{r}
city_search_frequency_loans <- loans_data2 %>%
  group_by(location) %>%
  summarize(total_hits = sum(hits, na.rm = TRUE)) %>%
  arrange(desc(total_hits))

city_search_frequency_loans
```
focus on the interest_by_city part
```{r}
crime_data2 <- res_hw1$interest_by_city[which(res_hw1$interest_by_city$keyword == "crime"), ]

```

Group the data about the "crime" keyword by city
Then look at the top cities
```{r}
city_search_frequency_crime <- crime_data2 %>%
  group_by(location) %>%
  summarize(total_hits = sum(hits, na.rm = TRUE)) %>%
  arrange(desc(total_hits))

city_search_frequency_crime
```

merge the loans and crime data
IN THIS X IS LOANS AND Y IS CRIME
```{r}
merged_data_crimeloans <- inner_join(city_search_frequency_loans, city_search_frequency_crime, by = "location")
head(merged_data_crimeloans)
```

Now merge it with the ACS Data
```{r}
merged_data_crimeloans_ACS <- inner_join(merged_data_crimeloans, acs_with_location, by = "location")
head(merged_data_crimeloans_ACS)
```

Compute the mean of the search popularity for both keywords for cities that have an above average median household income and for those that have an below average median household income. When building your pipe, start with creating the grouping variable and then proceed with the remaining tasks. What conclusions might you draw from this?

Calculate the average household income
```{r}
library(dplyr)
average_income_crimeloans_ACS <- mean(merged_data_crimeloans_ACS$hh_income, na.rm = TRUE)
average_income_crimeloans_ACS
```

create grouping variable
```{r}
merged_data_crimeloans_ACS <- merged_data_crimeloans_ACS %>%
  mutate(income_group = ifelse(hh_income > average_income_crimeloans_ACS, "Above Average", "Below Average"))
head(merged_data_crimeloans_ACS)
merged_data_crimeloans_ACS
```

group by income group
```{r}
grouped_data_crimeloans_ACS <- merged_data_crimeloans_ACS %>%
  group_by(income_group)
```

calculate the mean for both groups
```{r}
result_crimeloans_ACS <- grouped_data_crimeloans_ACS %>%
  summarize(
    mean_popularity_crime = mean(total_hits.y, na.rm = TRUE), #crime
    mean_popularity_loans = mean(total_hits.x, na.rm = TRUE) #loans
  )
print(result_crimeloans_ACS)
```

Is there a relationship between the median household income and the search popularity of the Google trends terms? Describe the relationship and use a scatterplot with qplot().

```{r}
library(ggplot2)
scatterplot_crimeloans_ACS <- ggplot(merged_data_crimeloans_ACS, aes(x = hh_income)) +
  geom_point(aes(y = total_hits.y, color = income_group), size = 3, alpha = 0.7) +
  geom_point(aes(y = total_hits.x, color = income_group), size = 3, shape = 2, alpha = 0.7) +
  labs(
    x = "Median Household Income",
    y = "Search Popularity",
    color = "Income Group",
    shape = "Income Group"
  ) +
  scale_shape_manual(values = c(19, 2)) +
  ggtitle("Relationship Between Median Household Income and Search Popularity") +
  theme_minimal()
print(scatterplot_crimeloans_ACS)
```
circles are crime, triangles are loans


Is there a relationship? look at the correlation
```{r}
correlation_crime_income <- cor(merged_data_crimeloans_ACS$total_hits.y, merged_data_crimeloans_ACS$hh_income, use = "complete.obs")
cat("Correlation between 'crime' and median household income:", correlation_crime_income, "\n")

correlation_loans_income <- cor(merged_data_crimeloans_ACS$total_hits.x, merged_data_crimeloans_ACS$hh_income, use = "complete.obs")
cat("Correlation between 'loans' and median household income:", correlation_loans_income, "\n")
```


#### last part ####

merge the mask and symptoms data
IN THIS X IS MASK AND Y IS SYMPTOMS
```{r}
city_search_frequency_mask # this is x
city_search_frequency_symptoms #this is y
merged_data_masksymptoms <- inner_join(city_search_frequency_mask, city_search_frequency_symptoms, by = "location")
merged_data_masksymptoms
```

Now merge it with the ACS Data
```{r}
merged_data_masksymptoms_ACS <- inner_join(merged_data_masksymptoms, acs_with_location, by = "location")
head(merged_data_masksymptoms_ACS)
```

Compute the mean of the search popularity for both keywords for cities that have an above average median household income and for those that have an below average median household income. When building your pipe, start with creating the grouping variable and then proceed with the remaining tasks. What conclusions might you draw from this?

Calculate the average household income
```{r}
library(dplyr)
average_income_masksymptoms_ACS <- mean(merged_data_masksymptoms_ACS$hh_income, na.rm = TRUE)
average_income_masksymptoms_ACS
```

create grouping variable
```{r}
merged_data_masksymptoms_ACS <- merged_data_masksymptoms_ACS %>%
  mutate(income_group = ifelse(hh_income > average_income_masksymptoms_ACS, "Above Average", "Below Average"))
head(merged_data_masksymptoms_ACS)
merged_data_masksymptoms_ACS
```

group by income group
```{r}
grouped_data_masksymptoms_ACS <- merged_data_masksymptoms_ACS %>%
  group_by(income_group)
```

calculate the mean for both groups
```{r}
result_masksymptoms_ACS <- grouped_data_masksymptoms_ACS %>%
  summarize(
    mean_popularity_symptoms = mean(total_hits.y, na.rm = TRUE), #symptoms
    mean_popularity_mask = mean(total_hits.x, na.rm = TRUE) #mask
  )
print(result_masksymptoms_ACS)
```

Is there a relationship between the median household income and the search popularity of the Google trends terms? Describe the relationship and use a scatterplot with qplot().

```{r}
library(ggplot2)
scatterplot_masksymptoms_ACS <- ggplot(merged_data_masksymptoms_ACS, aes(x = hh_income)) +
  geom_point(aes(y = total_hits.y, color = income_group), size = 3, alpha = 0.7) +
  geom_point(aes(y = total_hits.x, color = income_group), size = 3, shape = 2, alpha = 0.7) +
  labs(
    x = "Median Household Income",
    y = "Search Popularity",
    color = "Income Group",
    shape = "Income Group"
  ) +
  scale_shape_manual(values = c(19, 2)) +
  ggtitle("Relationship Between Median Household Income and Search Popularity") +
  theme_minimal()
print(scatterplot_masksymptoms_ACS)
```
circles are symptoms, triangles are mask

Is there a relationship? look at the correlation
```{r}
correlation_symptoms_income <- cor(merged_data_masksymptoms_ACS$total_hits.y, merged_data_masksymptoms_ACS$hh_income, use = "complete.obs")
cat("Correlation between 'symptoms' and median household income:", correlation_symptoms_income, "\n")

correlation_mask_income <- cor(merged_data_masksymptoms_ACS$total_hits.x, merged_data_masksymptoms_ACS$hh_income, use = "complete.obs")
cat("Correlation between 'mask' and median household income:", correlation_mask_income, "\n")
```


