---
title: "Assignment 4 Wednesday"
author: "Rowan Muhl"
date: "2023-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an individual assignment. Turn in this assignment as an HTML or PDF file to ELMS. Make sure to include the R Markdown or Quarto file that was used to generate it. Include the GitHub link for the repository containing these files.

Link to GitHub: https://github.com/kglemaud/glemaud-muhl-pistner-a1

```{r}
library(tidyverse)
library(DBI)
library(dbplyr)
library(bigrquery)
```

In this notebook we will use Google BigQuery, "Google's fully managed, petabyte scale, low cost analytics data warehouse". Some instruction on how to connect to Google BigQuery can be found here: https://db.rstudio.com/databases/big-query/.

You will need to set up a Google account with a project to be able to use this service. We will be using a public dataset that comes with 1 TB/mo of free processing on Google BigQuery. As long as you do not repeat the work in this notebook constantly, you should be fine with just the free tier.

Go to https://console.cloud.google.com and make sure you are logged in a non-university Google account. This may not work on a university G Suite account because of restrictions on those accounts. Create a new project by navigating to the dropdown menu at the top (it might say "Select a project") and selecting "New Project" in the window that pops up. Name it something useful.

After you have initialized a project, paste your project ID into the following chunk.

```{r}
project_id <- "surv-727-project-muhl"
```

We will connect to a public database, the Chicago crime database, which has data on crime in Chicago.

```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  dataset = "chicago_crime",
  billing = project_id
)
con
```

We can look at the available tables in this database using dbListTables.

Note: When you run this code, you will be sent to a browser and have to give Google permissions to Tidyverse API Packages. Make sure you select all to give access or else your code will not run.

```{r}
dbListTables(con)
```

Information on the `crime` table can be found here:

https://cloud.google.com/bigquery/public-data/chicago-crime-data

Write a first query that counts the number of rows of the `crime` table in the year 2016. Use code chunks with {sql connection = con} in order to write SQL code within the document.
```{r}
sql_query <- "SELECT COUNT(*) AS row_count FROM `bigquery-public-data.chicago_crime.crime` WHERE EXTRACT(YEAR FROM date) = 2016"
result <- dbGetQuery(con, sql_query)
result
```
269,841


Next, count the number of arrests grouped by primary_type in 2016. Note that is a somewhat similar task as above, with some adjustments on which rows should be considered. Sort the results, i.e. list the number of arrests in a descending order.
```{r}
sql_query2 <- "
SELECT primary_type, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE EXTRACT(YEAR FROM date) = 2016
AND arrest = TRUE
GROUP BY primary_type
ORDER BY arrest_count DESC
"

arrests_by_primary_type <- dbGetQuery(con, sql_query2)

arrests_by_primary_type
```

We can also use the date for grouping. Count the number of arrests grouped by hour of the day in 2016. You can extract the latter information from date via EXTRACT(HOUR FROM date). Which time of the day is associated with the most arrests?
```{r}
sql_query3 <- "
SELECT EXTRACT(HOUR FROM date) AS hour_of_day, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE EXTRACT(YEAR FROM date) = 2016
AND arrest = TRUE
GROUP BY hour_of_day
ORDER BY arrest_count DESC
"

arrests_by_hour_of_day <- dbGetQuery(con, sql_query3)

arrests_by_hour_of_day
```
The hour of the day associated with the most arrests is 10.


Focus only on HOMICIDE and count the number of arrests for this incident type, grouped by year. List the results in descending order.
```{r}
sql_query4 <- "
SELECT EXTRACT(YEAR FROM date) AS year, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE primary_type = 'HOMICIDE'
AND arrest = TRUE
GROUP BY year
ORDER BY year DESC
"

homicide_arrests_by_year <- dbGetQuery(con, sql_query4)

homicide_arrests_by_year
```

Find out which districts have the highest numbers of arrests in 2015 and 2016. That is, count the number of arrests in 2015 and 2016, grouped by year and district. List the results in descending order.
```{r}
sql_query5 <- "
SELECT EXTRACT(YEAR FROM date) AS year, district, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE EXTRACT(YEAR FROM date) IN (2015, 2016)
AND arrest = TRUE
GROUP BY year, district
ORDER BY year DESC, arrest_count DESC
"

arrests_by_year_and_district <- dbGetQuery(con, sql_query5)

arrests_by_year_and_district
```
In 2016, the districts with the highest numbers of arrests are 11, 7, and 6 (in that order).
In 2015, the districts with the highest numbers of arrests are 11, 7, and 15 (in that order).


Lets switch to writing queries from within R via the DBI package. Create a query object that counts the number of arrests grouped by primary_type of district 11 in year 2016. The results should be displayed in descending order. Execute the query.
```{r}
sql_query6 <- "
SELECT primary_type, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE EXTRACT(YEAR FROM date) = 2016
AND district = 11
AND arrest = TRUE
GROUP BY primary_type
ORDER BY arrest_count DESC
"

query6 <- dbSendQuery(con, sql_query6)

results <- dbFetch(query6)

dbClearResult(query6)

results
```

Try to write the very same query, now using the dbplyr package. For this, you need to first map the crime table to a tibble object in R.
```{r}
library(dplyr)
library(dbplyr)
library(DBI)


crime_tibble <- tbl(con,"crime")

query_dbplyr <- crime_tibble %>% 
  filter(year == 2016, district == 11) %>%
  group_by(primary_type) %>%
  summarize(arrest_count = n()) %>%
  arrange(desc(arrest_count))
print(query_dbplyr)



```


Again, count the number of arrests grouped by primary_type of district 11 in year 2016, now using dplyr syntax.
```{r}


library(dplyr)


result <- crime_tibble %>%
  filter(district == 11, year == 2016, arrest) %>%
  group_by(primary_type) %>%
  summarise(arrest_count = n()) %>%
  arrange(desc(arrest_count))


print(result)

```

Count the number of arrests grouped by primary_type and year, still only for district 11. Arrange the result by year.
```{r}
result <- crime_tibble %>%
  filter(district == 11, arrest) %>%
  group_by(primary_type, year) %>%
  summarise(arrest_count = n()) %>%
  arrange(year)
print(result)
```

Assign the results of the query above to a local R object.
```{r}
local_r_object <- result
```

Confirm that you pulled the data to the local environment by displaying the first ten rows of the saved data set.
```{r}
head(local_r_object, n = 10)
```

Close the connection.
```{r}
dbDisconnect(con)
```

