---
title: "homework 4"
author: "Rowan Muhl"
date: "2023-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1.1 Optimal matching, Mahalanobis distance, matrix multiplication
```{r}
# Covariance matrix
cov_matrix <- matrix(c(10, 0.2, 0.2, 1), 2, 2)
solve(cov_matrix)
```

```{r}
# Covariance matrix
cov_matrix <- matrix(c(10, 0.2, 0.2, 1), 2, 2)

# Covariates for treated units
X1 <- c(25, 1)  # Treated unit i=1
X2 <- c(30, 1)  # Treated unit i=2

# Control units' covariates
control_units <- matrix(c(30, 1, 40, 1), 2, 2, byrow = TRUE)

# Function to calculate Mahalanobis distance
mahalanobis_distance <- function(X, Y, cov_matrix) {
  diff <- X - Y
  distance <- sqrt(t(diff) %*% solve(cov_matrix) %*% diff)
  return(distance)
}

# Calculate Mahalanobis distances for each treated unit with all control units
distances1 <- apply(control_units, 2, function(j) mahalanobis_distance(X1, j, cov_matrix))
distances2 <- apply(control_units, 2, function(j) mahalanobis_distance(X2, j, cov_matrix))

# Find the matched control unit for each treated unit
matched_control1 <- which.min(distances1)
matched_control2 <- which.min(distances2)

# Create the table
table_data <- data.frame(
  "Matching pairs" = 1:2,
  "Treated i" = c("i=1", "i=2"),
  "Control j(i)" = c(matched_control1, matched_control2)
)

print(table_data)

```
To fill in the empty column, the first box is 2 and the second box is 2.


1.2 Estimate ACE using the matched pair
```{r}
# Covariates for treated units
X1 <- c(25, 1)  # Treated unit i=1
X2 <- c(30, 1)  # Treated unit i=2

# Control units' covariates
control_units <- matrix(c(30, 1, 40, 1), 2, 2, byrow = TRUE)

# Income for treated units
Income_treated <- c(15, 22)

# Income for matched control units (based on the previous matching)
Income_control <- c(10, 15)

# Calculate the Average Treatment Effect (ATE)
ACE <- mean(Income_treated - Income_control)

# Print the ATE
print(paste("Average Causal Effect (ACE):", ACE))

```
It's 6

1.3 Propensity score weights
```{r}
# Given data
e_x <- c(0.25, 4, 0.33, 0.5)
Income <- c(15, 22, 10, 15)

# Initialize vectors for PS weight and Income * weight
PS_weight <- rep(NA, length(e_x))
Income_weighted <- rep(NA, length(e_x))

# Calculate PS weight and Income * weight
for (i in 1:length(e_x)) {
  if (i <= 2) {
    PS_weight[i] <- 1 / e_x[i]  # Treated units
  } else {
    PS_weight[i] <- 1 / (1 - e_x[i])  # Control units
  }
  
  Income_weighted[i] <- Income[i] * PS_weight[i]
}

# Create the table
table_data <- data.frame(
  "Treated i" = c("i=1", "i=2", "Control j=1", "Control j=2"),
  "e(x)" = e_x,
  "Income (=$10k)" = Income,
  "PS weight (w)" = PS_weight,
  "Income*w" = Income_weighted
)

print(table_data)

```
The values for the propensity score weights are 4, 0.25, 1.49, and 2.
The values for Income*w  are 60, 5.5, 14.93, and 30.

1.4 Average causal effect using the measure of risk difference (RD)
```{r}
# Given data
Income <- c(15, 22, 10, 15)
PS_weight <- c(1/0.25, 1/4, 1/0.33, 1/0.5)  # Assuming inverse of e(x) as PS weights

# Calculate Risk Difference
RD <- sum((Income[1:2] - Income[3:4]) * PS_weight[1:2]) / length(Income)

print(paste("Estimated Risk Difference (RD):", RD))

```
RD = 5.4375

2. prove that the weighted mean of untreated units equals the mean of potential outcome

Yi = Ai * Yi1 + (1 - Ai) * Yi0

E[(1 - Ai)wiYi] = E[(1 - Ai)wi(AiYi1 + (1 - Ai)Yi0)]

E[(1 - Ai)wi(AiYi1)] + E[(1 - Ai)wi((1 - Ai)Yi0)]

E[(1 - Ai)wi(AiYi1)] = E[0] = 0

E[(1 - Ai)wi((1 - Ai)Yi0)]

for (Ai = 0), this becomes E[wiYi0]

Which means that the weighted mean of the untreated units equals the mean of the potential outcome under A = 0

E[(1 - Ai)wiYi] = E[wiYi0] = E(Yi0)



3.a causal RD, RR, and OR by standardization method
```{r}
# Given data
data <- matrix(c(108, 252, 360, 24, 16, 40, 20, 30, 50, 40, 10, 50), nrow = 3, byrow = TRUE)
colnames(data) <- c("Asub0 = 1", "Asub0 = 0", "Asub0 = 1", "Asub0 = 0")
rownames(data) <- c("Y = 1", "Y = 0", "Total")

# Calculate probabilities
P_Y1_A1 <- data["Y = 1", "Asub0 = 1"] / data["Total", "Asub0 = 1"]
P_Y1_A0 <- data["Y = 1", "Asub0 = 0"] / data["Total", "Asub0 = 0"]

# Causal RD
RD <- P_Y1_A1 - P_Y1_A0

# Causal RR
RR <- P_Y1_A1 / P_Y1_A0

# Causal OR
OR <- (P_Y1_A1 / (1 - P_Y1_A1)) / (P_Y1_A0 / (1 - P_Y1_A0))

# Print the results
print(paste("Causal RD:", RD))
print(paste("Causal RR:", RR))
print(paste("Causal OR:", OR))

```
Causal RD = -4.14
Causal RR = 0.343
Causal OR = 1.57

3.b create weights and stabalized weights for MSM estimation
```{r}
# Given data
data <- matrix(c(108, 252, 360, 24, 16, 40, 20, 30, 50, 40, 10, 50), nrow = 3, byrow = TRUE)
colnames(data) <- c("Asub0 = 1", "Asub0 = 0", "Asub0 = 1", "Asub0 = 0")
rownames(data) <- c("Y = 1", "Y = 0", "Total")

# Calculate probabilities
P_A1 <- data["Total", "Asub0 = 1"] / sum(data["Total", c("Asub0 = 1", "Asub0 = 0")])
P_A0_given_L0 <- data["Y = 0", "Asub0 = 0"] / data["Total", "Asub0 = 0"]

# Standard weights
SW <- 1 / P_A1

# Stabilized weights
SW_star <- 1 / (P_A1 * P_A0_given_L0)

# Print the results
print("Standard Weights:")
print(SW)

print("Stabilized Weights:")
print(SW_star)

```


3.c Esitimate the causal RD, RR, and OR using MSM method with the model Y=A where the outcome Y is specified as a binary variable

```{r}
#install.packages("survey")
#install.packages("gee")
library(gee)
```

```{r}
# Given data
data <- matrix(c(108, 252, 360, 24, 16, 40, 20, 30, 50, 40, 10, 50), nrow = 3, byrow = TRUE)
colnames(data) <- c("Asub0 = 1", "Asub0 = 0", "Asub0 = 1", "Asub0 = 0")
rownames(data) <- c("Y = 1", "Y = 0", "Total")

# Extract the relevant data
Y_A1 <- c(data["Y = 1", "Asub0 = 1"], data["Y = 0", "Asub0 = 1"])
Y_A0 <- c(data["Y = 1", "Asub0 = 0"], data["Y = 0", "Asub0 = 0"])
Total_A1 <- sum(data["Total", "Asub0 = 1"])
Total_A0 <- sum(data["Total", "Asub0 = 0"])

# Calculate proportions
prop_Y_A1 <- Y_A1 / Total_A1
prop_Y_A0 <- Y_A0 / Total_A0

# Calculate risk difference, risk ratio, and odds ratio
causal_RD <- prop_Y_A1[1] - prop_Y_A0[1]
causal_RR <- prop_Y_A1[1] / prop_Y_A0[1]
causal_OR <- (prop_Y_A1[1] / (1 - prop_Y_A1[1])) / (prop_Y_A0[1] / (1 - prop_Y_A0[1]))

# Calculate standard errors
se_RD <- sqrt((prop_Y_A1[1] * (1 - prop_Y_A1[1]) / Total_A1) + (prop_Y_A0[1] * (1 - prop_Y_A0[1]) / Total_A0))
se_RR <- sqrt((1 / (prop_Y_A0[1] * Total_A0)) + (1 / (prop_Y_A1[1] * Total_A1)))
se_OR <- sqrt((1 / (prop_Y_A0[1] * (1 - prop_Y_A0[1]))) + (1 / (prop_Y_A1[1] * (1 - prop_Y_A1[1]))))

# Calculate confidence intervals
ci_RD <- c(causal_RD - 1.96 * se_RD, causal_RD + 1.96 * se_RD)
ci_RR <- c(causal_RR * exp(-1.96 * se_RR), causal_RR * exp(1.96 * se_RR))
ci_OR <- c(causal_OR * exp(-1.96 * se_OR), causal_OR * exp(1.96 * se_OR))

# Print the results
print(paste("Causal RD:", causal_RD, " (", ci_RD[1], ", ", ci_RD[2], ")"))
print(paste("Causal RR:", causal_RR, " (", ci_RR[1], ", ", ci_RR[2], ")"))
print(paste("Causal OR:", causal_OR, " (", ci_OR[1], ", ", ci_OR[2], ")"))

```
Causal RD = -4.14
Causal RR = 0.343
Causal OR = 1.57
